% ************************** Thesis Acknowledgements *****************************

\begin{acknowledgements}      

Behind this dissertation, is an battalion, who either inspired, nudged, guided, mentored or supported me, to eventually pursue something which had been a dream in making for almost 20 years. 

Firstly, I have to express my deepest gratitude to my advisor and mentor Dr. Nishanth Sastry. I still remember the long and detailed email he wrote, even before I joined his group, unwrapping every doubt I had about all things Ph.D. It was because of this email and the brutally honest but empathic advice within, is why I decided to take a leap of faith and pursue this journey under his wing.  At every step, he has inspired me to explore the difficult and be fearless doing so. With a deliberate balance of control and freedom, his style of mentoring is something I am truly thankful for and wish to internalize at some point.

The biggest curse of doctoral studies is the loneliness that it breeds. But I was immensely lucky to find friends on and off campus. I am truly grateful to find friends like Aravindh Raman, Changtao Zhong, Dmytro Karamshuk, Pushkal Agarwal, Emeka Obiedu and Miriam Redi, with whom I have had the pleasure of sharing countless coffees and endless inspiring chats. Some of these chats have even blossomed into full fledged research projects and generated grants and/or publications. I have been blessed with friends like Anand, Sanika, Marc and Roxy, who filled our life with fun, love and a warm feeling of being part of a family. 

I have been fortunate in life to find mentors at different stages, who have shaped my career and life in some way. Here too, I was lucky to find mentors either through collaborations or shared interests, who have been quintessential in shaping my Ph.D. and my career. I would like to thank Luca Aiello, Miriam Redi, Gareth Tyson, Daniele Quercia, Pietro Panzarasa and Anna Simoni for their support and valuable mentoring. 

Finally, it only makes sense to connect the dots in hindsight. To that end, I have to thank my friends and colleagues from Santa Barbara who gave me honest and balanced advice when I needed the most. I would like to thank Advait, Geetak, Bugra and Gina, who made sure my desire to do a Ph.D. became a drive and resulted in an action. 

My decision to pursue a Ph.D., could not have been possible without the unconditional support of my family. Every curious thought in my mind can be traced back to how I was raised. My mother Medha and my sister Chanda, have been the biggest driving force behind all my actions. They have always inspired me to follow my heart, even if at times it meant me going away from them for a long haul. I would like to thank my other half of the family, Sanjeev, Sandhya and Rishikesh, who had faith in me when I needed it the most.

Last, but by long shot not the least, I have to thank my wife Geetika, who has been like a light house in a turbulent sea. Unwavering in her support and unparalleled in her guidance, she has gone above an beyond, to make this dissertation a reality. And hence I believe it is as much hers, as mine. And to conclude, I would like to thank Arya, our new born daughter. She made sure that I keep to the schedule of this Ph.D. by announcing hers when I started writing up. 

This has been one bumpy uphill ride, and as it comes towards an end I realise that everything looks beautiful from up here.
\end{acknowledgements}
